Distributional Regression Forests for Probabilistic Precipitation Forecasting in Complex Terrain

### 2018-01-04 pkg/disttree/inst/draft/draft.Rnw
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1902 255f4249-2263-4417-b1ea-8498779e27a7
Former-commit-id: 58d1788f010af26e5f2bfa24b247f6e9b58a4305
small corrections and additional references
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1919 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: b4519e640ab265fb8ea81a1915a085e5a036d9c7

  if(type == "exp"){
    rmse <- NULL
    colnames <- NULL
    col <- NULL
    legendnames <- NULL
    if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.dt"], simres$resmat[,"av.rmse.exp.obs.dt"])
      colnames <- c(colnames, "dt.true", "dt.obs")
      col <- c(col, pal["tree"])
      legendnames <- c(legendnames, "tree")
    }
    if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.df"], simres$resmat[,"av.rmse.exp.obs.df"])
      colnames <- c(colnames, "df.true", "df.obs")
      col <- c(col, pal["forest"])
      legendnames <- c(legendnames, "forest")
    }
    if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.g"], simres$resmat[,"av.rmse.exp.obs.g"])
      colnames <- c(colnames, "g.true", "g.obs")
      col <- c(col, pal["gamlss"])
      legendnames <- c(legendnames, "gamlss")
    }
    if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.b"], simres$resmat[,"av.rmse.exp.obs.b"])
      colnames <- c(colnames, "b.true", "b.obs")
      col <- c(col, pal["bamlss"])
      legendnames <- c(legendnames, "bamlss")
    }
    if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.gb"], simres$resmat[,"av.rmse.exp.obs.gb"])
      colnames <- c(colnames, "gb.true", "gb.obs")
      col <- c(col, pal["gamboostLSS"])
      legendnames <- c(legendnames, "gamboostLSS")
    }
    if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.rf"], simres$resmat[,"av.rmse.exp.obs.rf"])
      colnames <- c(colnames, "rf.true", "rf.obs")
      col <- c(col, pal["randomForest"])
      legendnames <- c(legendnames, "randomForest")
    }
    if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.cf"], simres$resmat[,"av.rmse.exp.obs.cf"])
      colnames <- c(colnames, "cf.true","cf.obs")
      col <- c(col, pal["cforest"])
      legendnames <- c(legendnames, "cforest")
    }

@@

    rmse <- NULL
    colnames <- NULL
    col <- NULL
    legendnames <- NULL
    if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.dt"], simres$resmat[,"av.rmse.sigma.dt"])
      colnames <- c(colnames, "dt.true", "dt.obs")
      col <- c(col, pal["tree"])
      legendnames <- c(legendnames, "tree")
    }
    if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.df"], simres$resmat[,"av.rmse.sigma.df"])
      colnames <- c(colnames, "df.true", "df.obs")
      col <- c(col, pal["forest"])
      legendnames <- c(legendnames, "forest")
    }
    if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.g"], simres$resmat[,"av.rmse.sigma.g"])
      colnames <- c(colnames, "g.true", "g.obs")
      col <- c(col, pal["gamlss"])
      legendnames <- c(legendnames, "gamlss")
    }
    if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.b"], simres$resmat[,"av.rmse.sigma.b"])
      colnames <- c(colnames, "b.true", "b.obs")
      col <- c(col, pal["bamlss"])
      legendnames <- c(legendnames, "bamlss")
    }
    if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.gb"], simres$resmat[,"av.rmse.sigma.gb"])
      colnames <- c(colnames, "gb.true", "gb.obs")
      col <- c(col, pal["gamboostLSS"])
      legendnames <- c(legendnames, "gamboostLSS")
    }
    if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.rf"], simres$resmat[,"av.rmse.sigma.rf"])
      colnames <- c(colnames, "rf.true", "rf.obs")
      col <- c(col, pal["randomForest"])
      legendnames <- c(legendnames, "randomForest")
    }
    if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.cf"], simres$resmat[,"av.rmse.sigma.cf"])
      colnames <- c(colnames, "cf.true","cf.obs")
      col <- c(col, pal["cforest"])
      legendnames <- c(legendnames, "cforest")
    }

@@

  ll <- NULL
  colnames <- NULL
  col <- NULL
  legendnames <- NULL
  if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.dt"])
    colnames <- c(colnames, "dt.ll")
    col <- c(col, pal["tree"])
    legendnames <- c(legendnames, "tree")
  }
  if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.df"])
    colnames <- c(colnames, "df.ll")
    col <- c(col, pal["forest"])
    legendnames <- c(legendnames, "forest")
  }
  if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.g"])
    colnames <- c(colnames, "g.true")
    col <- c(col, pal["gamlss"])
    legendnames <- c(legendnames, "gamlss")
  }
  if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.b"])
    colnames <- c(colnames, "b.true")
    col <- c(col, pal["bamlss"])
    legendnames <- c(legendnames, "bamlss")
  }
  if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.gb"])
    colnames <- c(colnames, "gb.true")
    col <- c(col, pal["gamboostLSS"])
    legendnames <- c(legendnames, "gamboostLSS")
  }
  if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.rf"])
    colnames <- c(colnames, "rf.true")
    col <- c(col, pal["randomForest"])
    legendnames <- c(legendnames, "randomForest")
  }
  if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.cf"])
    colnames <- c(colnames, "cf.true")
    col <- c(col, pal["cforest"])
    legendnames <- c(legendnames, "cforest")
  }

@@

  crps <- NULL
  colnames <- NULL
  col <- NULL
  legendnames <- NULL
  if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.dt"])
    colnames <- c(colnames, "dt.crps")
    col <- c(col, pal["tree"])
    legendnames <- c(legendnames, "tree")
  }
  if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.df"])
    colnames <- c(colnames, "df.crps")
    col <- c(col, pal["forest"])
    legendnames <- c(legendnames, "forest")
  }
  if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.g"])
    colnames <- c(colnames, "g.true")
    col <- c(col, pal["gamlss"])
    legendnames <- c(legendnames, "gamlss")
  }
  if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.b"])
    colnames <- c(colnames, "b.true")
    col <- c(col, pal["bamlss"])
    legendnames <- c(legendnames, "bamlss")
  }
  if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.gb"])
    colnames <- c(colnames, "gb.true")
    col <- c(col, pal["gamboostLSS"])
    legendnames <- c(legendnames, "gamboostLSS")
  }
  if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.rf"])
    colnames <- c(colnames, "rf.true")
    col <- c(col, pal["randomForest"])
    legendnames <- c(legendnames, "randomForest")
  }
  if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.cf"])
    colnames <- c(colnames, "cf.true")
    col <- c(col, pal["cforest"])
    legendnames <- c(legendnames, "cforest")
  }

@@

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale and shape (GAMLSS, \citealt{stasinopoulos2005}) each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees (\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). 
%%%<TH> no references in abstract </TH>
This class of models 
%%%<TH> "this" means trees and forests? </TH>
will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.

@@

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/lisa-schlosser.html}, \url{http://retostauffer.org/}, \url{https://eeecon.uibk.ac.at/~zeileis/}\\

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
In regression analysis various models have been developed to describe the relationship between a response variable and a (set of) covariate(s). Starting from linear models (LM) being extended to generalized linear models (GLM) the limitation to model only linear effects has lead to the development of generalized additive models (GAM) also capturing non-linear effects as illustrated in Figure~\ref{Devel_parmod}. 
%%%<TH> more explantion for Fig 1 needed </TH>
These models consider only the expectation of the underlying distribution, however, in recent years there has been a paradigm shift from regression models for the mean to distributional regression approaches for probabilistic forecasting. For example in the field of precipitation forecasting linear models such as ensemble model output statistics (EMOS, \citealt{gneiting2005calibrated}) usually acting on the assumption of a normally distributed response variable allow for both parameters, the mean and the variance, to depend linearly on covariates. Based on the resulting model the probability of precipitation can be estimated as well as the expected amount of precipitation. Therefore, these models provide common and widely used tools, also due to their intuitive structure and interpretability. However, the restriction to model only linear effects has risen the interest in applying more flexible models. Generalized additive models for location, scale and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM. These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth and homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain in mountainous regions various factors can change rapidly and their effects are difficult to be represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree-structured models such as regression trees 
%%%<TH> citation Loh does not fit, maybe omit citations anyway </TH>
(\citealt{loh2011classification}) or random forests (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.

For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. Therefore, we are going to introduce a novel general distributional framework for these methods by presenting distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. The developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated in Figure~\ref{Devel_treeforest}.
%%% <TH> better description for Fig 2 </TH>

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a very large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a forest model are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values 
%%% well fitting predicted expected values -> appropriate probabilistic forecasts? </TH>
are provided by this model as they are close to the observed values. Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis. In section \ref{ProbabilisticForecasting} this example will be further discussed and it will be shown that the predictions made by the forest model are reasonably well fitting, not only for the mean parameter of the distribution but also for its second parameter, the scale parameter, and therefore the whole distribution.

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions of total precipitation at station Axams for July 24 in 2009, 2010, 2011 and 2012 by a distributional forest model learned on data from 1985--2008. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

@@

A common way of including these covariates is to consider them as regressors in a distributional model such that each parameter $\theta$ depends on covariates $\bold{X} = (X_1,\ldots,X_m)$. This could for example be done by a linear function $g$. In that way the fitted model would be specified as $D(Y, \theta = g(\bold{X}))$. In order to avoid the restriction to linear effects one might also use a set of smooth functions $\{f_i\}_{i = 1,\ldots,p}$ to specify a more flexible model $D(Y, \theta = f_1(\bold{X}) + \ldots + f_p(\bold{X}))$ applying a GAM structure. Another approach to distributional regression is proposed by \citealt{Klein2015} where each parameter of a possibly complex distribution can be modeled by an additive term consisting of a variety of different functional effect types. These include non-linear effects, spatial effects, random coefficients and interaction surfaces which extends the flexibility compared to parametric models such as GAMLSS modeling only smooth effects.

>>>>

  if(type == "exp"){
    rmse <- NULL
    colnames <- NULL
    col <- NULL
    legendnames <- NULL
    if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.dt"], simres$resmat[,"av.rmse.exp.obs.dt"])
      colnames <- c(colnames, "dt.true", "dt.obs")
      col <- c(col, pal["tree"])
      legendnames <- c(legendnames, "tree")
    }
    if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.df"], simres$resmat[,"av.rmse.exp.obs.df"])
      colnames <- c(colnames, "df.true", "df.obs")
      col <- c(col, pal["forest"])
      legendnames <- c(legendnames, "distributional forest")
    }
    if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.g"], simres$resmat[,"av.rmse.exp.obs.g"])
      colnames <- c(colnames, "g.true", "g.obs")
      col <- c(col, pal["gamlss"])
      legendnames <- c(legendnames, "gamlss")
    }
    if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.b"], simres$resmat[,"av.rmse.exp.obs.b"])
      colnames <- c(colnames, "b.true", "b.obs")
      col <- c(col, pal["bamlss"])
      legendnames <- c(legendnames, "bamlss")
    }
    if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.gb"], simres$resmat[,"av.rmse.exp.obs.gb"])
      colnames <- c(colnames, "gb.true", "gb.obs")
      col <- c(col, pal["gamboostLSS"])
      legendnames <- c(legendnames, "gamboostLSS")
    }
    if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.rf"], simres$resmat[,"av.rmse.exp.obs.rf"])
      colnames <- c(colnames, "rf.true", "rf.obs")
      col <- c(col, pal["randomForest"])
      legendnames <- c(legendnames, "randomForest")
    }
    if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.exp.true.cf"], simres$resmat[,"av.rmse.exp.obs.cf"])
      colnames <- c(colnames, "cf.true","cf.obs")
      col <- c(col, pal["cforest"])
      legendnames <- c(legendnames, "cforest")
    }

@@

    rmse <- NULL
    colnames <- NULL
    col <- NULL
    legendnames <- NULL
    if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.dt"], simres$resmat[,"av.rmse.sigma.dt"])
      colnames <- c(colnames, "dt.true", "dt.obs")
      col <- c(col, pal["tree"])
      legendnames <- c(legendnames, "tree")
    }
    if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.df"], simres$resmat[,"av.rmse.sigma.df"])
      colnames <- c(colnames, "df.true", "df.obs")
      col <- c(col, pal["forest"])
      legendnames <- c(legendnames, "distributional forest")
    }
    if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.g"], simres$resmat[,"av.rmse.sigma.g"])
      colnames <- c(colnames, "g.true", "g.obs")
      col <- c(col, pal["gamlss"])
      legendnames <- c(legendnames, "gamlss")
    }
    if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.b"], simres$resmat[,"av.rmse.sigma.b"])
      colnames <- c(colnames, "b.true", "b.obs")
      col <- c(col, pal["bamlss"])
      legendnames <- c(legendnames, "bamlss")
    }
    if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.gb"], simres$resmat[,"av.rmse.sigma.gb"])
      colnames <- c(colnames, "gb.true", "gb.obs")
      col <- c(col, pal["gamboostLSS"])
      legendnames <- c(legendnames, "gamboostLSS")
    }
    if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.rf"], simres$resmat[,"av.rmse.sigma.rf"])
      colnames <- c(colnames, "rf.true", "rf.obs")
      col <- c(col, pal["randomForest"])
      legendnames <- c(legendnames, "randomForest")
    }
    if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
      rmse <- cbind(rmse, simres$resmat[,"av.rmse.mu.cf"], simres$resmat[,"av.rmse.sigma.cf"])
      colnames <- c(colnames, "cf.true","cf.obs")
      col <- c(col, pal["cforest"])
      legendnames <- c(legendnames, "cforest")
    }

@@

  ll <- NULL
  colnames <- NULL
  col <- NULL
  legendnames <- NULL
  if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.dt"])
    colnames <- c(colnames, "dt.ll")
    col <- c(col, pal["tree"])
    legendnames <- c(legendnames, "tree")
  }
  if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.df"])
    colnames <- c(colnames, "df.ll")
    col <- c(col, pal["forest"])
    legendnames <- c(legendnames, "distributional forest")
  }
  if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.g"])
    colnames <- c(colnames, "g.true")
    col <- c(col, pal["gamlss"])
    legendnames <- c(legendnames, "gamlss")
  }
  if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.b"])
    colnames <- c(colnames, "b.true")
    col <- c(col, pal["bamlss"])
    legendnames <- c(legendnames, "bamlss")
  }
  if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.gb"])
    colnames <- c(colnames, "gb.true")
    col <- c(col, pal["gamboostLSS"])
    legendnames <- c(legendnames, "gamboostLSS")
  }
  if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.rf"])
    colnames <- c(colnames, "rf.true")
    col <- c(col, pal["randomForest"])
    legendnames <- c(legendnames, "randomForest")
  }
  if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
    ll <- cbind(ll, simres$resmat[,"av.loglik.cf"])
    colnames <- c(colnames, "cf.true")
    col <- c(col, pal["cforest"])
    legendnames <- c(legendnames, "cforest")
  }

@@

  crps <- NULL
  colnames <- NULL
  col <- NULL
  legendnames <- NULL
  if("av.rmse.exp.true.dt" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.dt"])
    colnames <- c(colnames, "dt.crps")
    col <- c(col, pal["tree"])
    legendnames <- c(legendnames, "tree")
  }
  if("av.rmse.exp.true.df" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.df"])
    colnames <- c(colnames, "df.crps")
    col <- c(col, pal["forest"])
    legendnames <- c(legendnames, "distributional forest")
  }
  if("av.rmse.exp.true.g" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.g"])
    colnames <- c(colnames, "g.true")
    col <- c(col, pal["gamlss"])
    legendnames <- c(legendnames, "gamlss")
  }
  if("av.rmse.exp.true.b" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.b"])
    colnames <- c(colnames, "b.true")
    col <- c(col, pal["bamlss"])
    legendnames <- c(legendnames, "bamlss")
  }
  if("av.rmse.exp.true.gb" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.gb"])
    colnames <- c(colnames, "gb.true")
    col <- c(col, pal["gamboostLSS"])
    legendnames <- c(legendnames, "gamboostLSS")
  }
  if("av.rmse.exp.true.rf" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.rf"])
    colnames <- c(colnames, "rf.true")
    col <- c(col, pal["randomForest"])
    legendnames <- c(legendnames, "randomForest")
  }
  if("av.rmse.exp.true.cf" %in% colnames(simres$resmat)) {
    crps <- cbind(crps, simres$resmat[,"av.crps.cf"])
    colnames <- c(colnames, "cf.true")
    col <- c(col, pal["cforest"])
    legendnames <- c(legendnames, "cforest")
  }

@@

\Abstract{
While many regression models consider only the mean of the response variable recent advances in parametric modeling have made it possible to model all distribution parameters. In that way the underlying distribution can be fully specified allowing for a wide range of analysis and prediction methods which are particularly valuable in applications such as probabilistic weather forecasting. In generalized additive models for location, scale and shape %(GAMLSS, \citealt{stasinopoulos2005}) 
each distribution parameter can be modeled separately by its own generalized additive model (GAM) capturing linear and non-linear effects. A different approach that would make it possible to also model non-additive effects and interactions is to apply tree-structured models such as regression trees %(\citealt{loh2011classification}) 
or random forests. %(\citealt{breiman2001random}). 
%%%<TH> "random forest" generisch; Breiman and Cutler's randomForest </TH>
This class of tree and forest models will now be integrated in probabilistic forecasting providing a novel general distributional framework for these methods. In particular, we are going to introduce distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. 
%%%<TH> break-up very long sentence </TH>
Applying these novel approaches to precipitation data in a mountainous region with a very large number of numerical weather prediction quantities shows one of their main advantages in selecting variables automatically and illustrates their strengths as their performance is compared to GAMLSS models that are specified based either on subject-matter knowledge and experience from meteorology or based on a computationally more demanding boosting approach.

@@

\Address{
  Lisa Schlosser, Reto Stauffer, Achim Zeileis \\
  Department of Statistics \\
  Faculty of Economics and Statistics \\
  Universit\"at Innsbruck \\
  Universit\"atsstr.~15 \\
  6020 Innsbruck, Austria \\
  E-mail: \email{Lisa.Schlosser@uibk.ac.at}, \email{Reto.Stauffer@uibk.ac.at}, \email{Achim.Zeileis@R-project.org} \\
  URL: \url{https://www.uibk.ac.at/statistics/personal/schlosser-lisa/}, \url{http://retostauffer.org/}, \url{https://eeecon.uibk.ac.at/~zeileis/}\\

@@

\SweaveOpts{eval = TRUE}
\section{Introduction}
In regression analysis various models have been developed to describe the relationship between a response variable and a (set of) covariate(s). Starting from linear models (LM) being extended to generalized linear models (GLM) the limitation to model only linear effects has lead to the development of generalized additive models (GAM) also capturing non-linear effects. In Figure~\ref{Devel_parmod} this development is illustrated for the case of one covariate. In the left picture the expected value of the response variable is modeled by a linear function in the covariate. In the middle a GAM is applied where a smooth function allows for more flexibility.

These models consider only the expectation of the underlying distribution, however, in recent years there has been a paradigm shift from regression models for the mean to distributional regression approaches for probabilistic forecasting. For example in the field of precipitation forecasting linear models such as ensemble model output statistics (EMOS, \citealt{gneiting2005calibrated}) usually acting on the assumption of a normally distributed response variable allow for both parameters, the mean and the variance, to depend linearly on covariates. Based on the resulting model the probability of precipitation can be estimated as well as the expected amount of precipitation. Therefore, these models provide common and widely used tools, also due to their intuitive structure and interpretability. However, the restriction to model only linear effects has risen the interest in applying more flexible models. Generalized additive models for location, scale and shape (GAMLSS, \citealt{stasinopoulos2005}) offer a framework in which each distribution parameter can be modeled separately by its own GAM as illustrated in the right plot of Figure~\ref{Devel_parmod}. These flexible parametric methods provide well fitting models when applied on precipitation data in a rather smooth and homogeneous setting, e.g. areas where no abrupt changes in meteorological/climatological events appear. However, for example in complex terrain in mountainous regions various factors can change rapidly and their effects are difficult to be represented by a smooth parametric model. In this situation a different approach which can deal with these abrupt changes would be to apply tree-structured models such as regression trees (\citealt{breiman1984classification})
%%%<TH> citation Loh does not fit, maybe omit citations anyway </TH>
%(\citealt{loh2011classification}) 
or random forests such as Breiman and Cutler's randomForest (\citealt{breiman2001random}). Moreover, these models are particularly advantageous in case of many covariates with unknown effects or interactions as the variable selection is done automatically within the model fitting process. Therefore, no additional effort or knowledge for the specification of the model is required in advance. However, even though their main advantage of being able to model non-additive effects and interactions without any necessary prespecification can play to their strength, regression trees are not an adequate choice if (additional) smooth effects appear as well. In this situation a random forest provides a good compromise as it combines a set of trees such that the steps imposed by the trees are smoothed out in case of smooth effects while abrupt changes are still captured by the tree structure of the model.

For these reasons, regression trees and particularly random forests provide valuable features which can contribute to improving models for probabilistic forecasting. Therefore, we are going to introduce a novel general distributional framework for these methods by presenting distributional trees and forests which embed the idea of flexible parametric modeling into the framework of regression trees and random forests. In Figure~\ref{Devel_treeforest} the developments in the field of tree-structured models together with the novel approaches of distributional trees and forests are illustrated for the case of one covariate. In the top left picture the step function modeling the mean of the response variable represents the division into three terminal subgroups by a regression tree. This is extended in the bottom left picture where a distributional tree models not only the mean but also the variance of the response variable by its own step function. The top middle picture shows that combining an ensemble of trees to a random forest can lead to smoother effects while the tree structure assures that abrupt changes are captured as well. Again, an extension of the random forest model considering only the mean is made in the top right picture. Here a distributional forest models the mean and the variance capturing abrupt changes but also smooth effects.

@@

After explaining the novel methodology in Section~\ref{Methodology} of this paper the focus will then be put on an application in the field of probabilistic forecasting in Section~\ref{ProbabilisticForecasting}. The new method is applied on precipitation data in a mountainous region with a very large number of numerical weather prediction quantities and the results are compared to those of other methods commonly used in this research area. In a first setting only one observation station is taken into consideration. The models are trained on data from 24 years and probabilistic predictions of the consecutive 4 years are validated. A brief outlook on the results of this setting is given by Figure~\ref{axams_onestation}. For the 24th of July of 2009, 2010, 2011 and 2012 the density functions of the distributions estimated by a distributional forest are plotted together with the corresponding observations. This illustrates that well fitting predicted expected values 
%%% well fitting predicted expected values -> appropriate probabilistic forecasts? </TH>
are provided by this model as they are close to the observed values. Beside the expected values the predictions provided by distributional models offer a wide range of information as they specify the whole distribution and therefore enable further methods of inference and analysis. In section \ref{ProbabilisticForecasting} this example will be further discussed and it will be shown that the predictions made by the distributional forest are reasonably well fitting, not only for the mean parameter of the distribution but also for its second parameter, the scale parameter, and therefore the whole distribution.

@@

# add labels
text(x = -0.8, y = lh1, labels = "2009", col = "red", cex = 0.8)
text(x = -0.8, y = lh2, labels = "2010", col = "blue", cex = 0.8)
text(x = -0.8, y = lh3, labels = "2011", col = "darkgreen", cex = 0.8)
text(x = -0.8, y = lh4, labels = "2012", col = "purple", cex = 0.8)
@
\caption{\label{axams_onestation}Predictions of total precipitation at station Axams for July 24 in 2009, 2010, 2011 and 2012 by a distributional forest learned on data from 1985--2008. Observations and predictions are left censored at 0. The corresponding point mass is shown at the censoring point (0).}
\end{center}
\end{figure}

@@

A common way of including these covariates is to consider them as regressors in a distributional model such that each parameter $\theta$ depends on covariates $\bold{Z} = (Z_1,\ldots,Z_m)$. This could for example be done by a linear function $g$. In that way the fitted model would be specified as $D(Y, \theta = g(\bold{Z}))$. In order to avoid the restriction to linear effects one might also use a set of smooth functions $\{f_i\}_{i = 1,\ldots,p}$ to specify a more flexible model $D(Y, \theta = f_1(\bold{Z}) + \ldots + f_p(\bold{Z}))$ applying a GAM structure. Another approach to distributional regression is proposed by \citealt{Klein2015} where each parameter of a possibly complex distribution can be modeled by an additive term consisting of a variety of different functional effect types. These include non-linear effects, spatial effects, random coefficients and interaction surfaces which extends the flexibility compared to parametric models such as GAMLSS modeling only smooth effects.
